Disease Knowledge Transfer across Alzheimer's Variants

### 2018-10-09 paper/paper.tex
updated figures, added quantification in paper

@@

\begin{figure}[h]
 \centering
 \includegraphics[width=0.8\textwidth,trim=0 0 0 80,clip]{figures/disease_knowledge_transfer.png}
 \caption{Outline of the proposed framework for joint modelling of multiple diseases. We assume that each disease can be modelled as the evolution of abstract dysfunctionality scores (Y-axis, top row), each one related to different brain regions. Each region-specific dysfunctionality score then further models (X-axis, bottom row) the progression of several modality-specific biomarkers within that same region. The biomarker correlations within the bottom units are assumed to be disease agnostic and shared across all diseases modelled. Disease knowledge transfer can then be achieved via the disease-agnostic units.}
 \label{fig:diagram}
\end{figure}

@@

\newcommand{\expFld}{figures}

% PCA vs tAD disease space
% \vspace{-2em}
\begin{picture}(5,5)
\put(0,-70){\textbf{\Large{A}}}
\end{picture}
\begin{figure}[H]
\centering
\begin{subfigure}{\textwidth}
\centering
\includegraphics[width=0.8\textwidth, trim=90 0 110 0, clip]{figures/unit1_allTraj_tad-drcTinyPen1_JMD.png} 
% \caption{}
% \label{fig:occipUnit}
\end{subfigure}

\begin{picture}(5,5)
\put(0,20){\textbf{\Large{B}}}
\end{picture}
\begin{subfigure}{0.47\textwidth}
\centering
% typical AD\\
\includegraphics[width=0.8\textwidth, trim=0 0 0 20, clip]{figures/tAD_trajSameSpace_tad-drcTinyPen1_JMD.png} 
% \caption{C}
\end{subfigure}
\begin{picture}(5,5)
\put(0,20){\textbf{\Large{C}}}
\end{picture}
\begin{subfigure}{0.47\textwidth}
\centering
% PCA\\
\includegraphics[width=0.8\textwidth, trim=0 0 0 20, clip]{figures/PCA_trajSameSpace_tad-drcTinyPen1_JMD.png} 
% \caption{PCA}
\end{subfigure}
\caption{(A) Estimated biomarker trajectories in the occipital functional unit. Subject data from ADNI and our local cohort are also shown. The X-axis, defined as the occipital dysfunctionality score, represents the time-shifts (in months) of each subject. Red lines represent samples from the trajectory posterior. The Y-axis measures biomarker values (normalised). (B-C) Progression of dysfunctionality scores for (B) typical AD and (C) PCA.}
\label{fig:pcaTadDisSpace}
\end{figure}

% estimated (hypothetical) trajectories in PCA: DTI, FDG, AV45, AV1451.Volumetric trajectories were based on PCA MRI data.
\begin{figure}
 \includegraphics[width=\textwidth, trim=0 20 0 0, clip]{figures/trajDisSpaceOverlap_PCA_tad-drcTinyPen1_JMD.png}
 \caption{Estimated trajectories for the PCA cohort. The only data that were available were the MRI volumetric data. The dynamics of the other biomarkers has been inferred by the model using data from typical AD, and taking into account the different spatial distribution of pathology in PCA as compared to typical AD.}
 \label{fig:PCAtrajByModality}
\end{figure}

@@

\begin{figure}
 \includegraphics[width=\textwidth, trim=0 0 0 0, clip]{figures/validDtiPCA.png}
 \caption{Validation of the Disease Knowledge Transfer Model using DTI data of PCA subjects from our local cohort. Each subject has been shifted along the X-axis only according to their MRI data. No DTI data from PCA subjects was used to fit the model.}
\label{fig:DTIvalid}
\end{figure}

>>>>

\newcommand{\expFld}{../resfiles/tad-drcTiny_JMD}
\newcommand{\jmdFld}{..}

@@

\begin{figure}[h]
 \centering
 \includegraphics[width=0.8\textwidth,trim=0 0 0 0,clip]{figures/disease_knowledge_transfer.pdf}
 \caption{Outline of the proposed framework for joint modelling of multiple diseases. We assume that each disease can be modelled as the evolution of abstract dysfunctionality scores (Y-axis, top row), each one related to different brain regions. Each region-specific dysfunctionality score then further models (X-axis, bottom row) the progression of several modality-specific biomarkers within that same region. The biomarker correlations within the bottom units are assumed to be disease agnostic and shared across all diseases modelled. Disease knowledge transfer can then be achieved via the disease-agnostic units.}
 \label{fig:diagram}
\end{figure}

@@

% PCA vs tAD disease space
% \vspace{-2em}
\begin{picture}(5,5)
\put(0,-70){\textbf{\Large{A}}}
\end{picture}
\begin{figure}[H]
\centering
\begin{subfigure}{\textwidth}
\centering
\includegraphics[width=0.8\textwidth, trim=90 0 110 0, clip]{\expFld/unit1_allTraj.png} 
% \caption{}
% \label{fig:occipUnit}
\end{subfigure}

\begin{picture}(5,5)
\put(0,20){\textbf{\Large{B}}}
\end{picture}
\begin{subfigure}{0.47\textwidth}
\centering
% typical AD\\
\includegraphics[width=0.8\textwidth, trim=0 0 0 20, clip]{\expFld/dis0_tAD_allTrajZeroOne.png} 
% \caption{C}
\end{subfigure}
\begin{picture}(5,5)
\put(0,20){\textbf{\Large{C}}}
\end{picture}
\begin{subfigure}{0.47\textwidth}
\centering
% PCA\\
\includegraphics[width=0.8\textwidth, trim=0 0 0 20, clip]{\expFld/dis1_PCA_allTrajZeroOne.png} 
% \caption{PCA}
\end{subfigure}
\caption{(A) Estimated biomarker trajectories in the occipital functional unit. Subject data from ADNI and our local cohort are also shown. The X-axis, defined as the occipital dysfunctionality score, represents the time-shifts (in months) of each subject. Red lines represent samples from the trajectory posterior. The Y-axis measures biomarker values (normalised). (B-C) Progression of dysfunctionality scores for (B) typical AD and (C) PCA.}
\label{fig:pcaTadDisSpace}
\end{figure}

% estimated (hypothetical) trajectories in PCA: DTI, FDG, AV45, AV1451.Volumetric trajectories were based on PCA MRI data.
\begin{figure}
 \includegraphics[width=\textwidth, trim=0 20 0 0, clip]{\expFld/trajDisSpaceOverlap_PCA_tad-drcTinyPen5_JMD.png}
 \caption{Estimated trajectories for the PCA cohort. The only data that were available were the MRI volumetric data. The dynamics of the other biomarkers has been inferred by the model using data from typical AD, and taking into account the different spatial distribution of pathology in PCA as compared to typical AD.}
 \label{fig:PCAtrajByModality}
\end{figure}

@@

\begin{table}
\begin{tabular}{c | c c c c c c}
\textbf{Model} & \textbf{Cingulate} & \textbf{Frontal} & \textbf{Hippocampus} & \textbf{Occipital} & \textbf{Parietal} & \textbf{Temporal}\\
& \multicolumn{6}{c}{\textbf{Prediction Error (MSE)}}\\
DKT & 0.09$\pm$0.04 & 0.03$\pm$0.01 & 0.18$\pm$0.03 & 0.04$\pm$0.02 & 0.06$\pm$0.02 & 0.04$\pm$0.02\\
Latent stage model & 0.09$\pm$0.04 & 0.03$\pm$0.01 & 0.17$\pm$0.03 & 0.04$\pm$0.02 & 0.06$\pm$0.02 & 0.04$\pm$0.02\\
Linear Model & 0.05$\pm$0.02* & 0.15$\pm$0.04* & 0.09$\pm$0.03* & 0.07$\pm$0.03* & 0.07$\pm$0.02* & 0.07$\pm$0.02*\\
& \multicolumn{6}{c}{\textbf{Rank Correlation (Spearman rho)}}\\
DKT  & 0.76  & 0.48  & 0.76  & 0.55  & 0.55  & 0.33 \\
Latent stage model  & 0.76  & 0.49  & 0.80*  & 0.56  & 0.51*  & 0.33 \\
Linear Model  & 0.48*  & 0.31*  & 0.64*  & 0.61*  & 0.57*  & 0.27* \\
\end{tabular}
\caption[Performance evaluation of DKT and two simpler models]{Performance evaluation of DKT and two simpler models: a latent stage disease progression model and a linear univariate model of DTI predicted from MRI. The top table shows the prediction error (MSE) while the bottom table shows the rank correlation measured by Spearman's rho. (*) Statistically significant difference between the current model performance and DKT, based on a two-tailed t-test, Bonferroni corrected.}
\label{sec:dktPerfMetrics}
\end{table}
